Energy and the CES production function

### 2015-08-14 Articles/Article_EcolEcon_2015/EcolEcon_2015.Rnw
No spaces in chunk names.

<<Fitted and Residual Graph Functions>>=
fittedGraphNoCSP <- function(data){
  fittedGraphData <- select(data, Year, Country, iGDP, fitted, fittedExp, flavor, energy, CSP, nest)
  fittedGraphData <- melt(fittedGraphData, 
                     measure.vars = c("iGDP", "fitted", "fittedExp"), 
                     value.name = "value", 
                     variable.name = "variable")
  fittedGraphData$label <- with(fittedGraphData, paste(variable, energy))
  # Now collapse some of the information. Don't need multiple versions of the exponential model, e.g.
  fittedGraphData <- filter(fittedGraphData, label=="iGDP With energy" | label=="fittedExp With energy" |
                       label == "fitted With energy" | label == "fitted Without energy")
  fittedGraphData$label <- plyr::mapvalues(fittedGraphData$label, 
                                      from = c("iGDP With energy", 
                                               "fittedExp With energy",
                                               "fitted With energy",
                                               "fitted Without energy"), 
                                      to = c("GDP", 
                                             "Reference model",
                                             "With energy",
                                             "Without energy"))
  # Relevel factors
  fittedGraphData$label <- relevelFactor(as.factor(fittedGraphData$label), 
                                         c("GDP", "Reference model", "Without energy", "With energy"))
  # Make (and return) the graph
  ggplot() +
    geom_line(data=fittedGraphData, mapping=aes(x=Year, y=value, group=label, linetype=label)) +
    scale_linetype_manual(values=c(1,4,2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

@@

The capital-labor Cobb-Douglas production function shown in Equation~\ref{eq:CD}
can be augmented to include an energy term:
%
\begin{equation} \label{eq:CDe}
  y = \theta A k^{\alpha_1} l^{\alpha_2} e^{\alpha_3} \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; ,
\end{equation}
%
where $e \equiv E/E_0$, and $E$ is in units of energy per time, typically TJ/year.
The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha_1 + \alpha_2 + \alpha_3 = 1$.
The term $A$ is known as total factor productivity,
and $\lambda$ is the Solow residual. 
%
<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models that reject the cost-share theorem. Solid line is historical GDP.">>=
fittedGraphNoCSP(filter(fitted_and_resid_data, model=="CD" & CSP=="Reject CSP"))
@

%
<<Graphs CD-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for Cobb-Douglas models.">>=
residualPlot(filter(fitted_and_resid_data, model=="CD" & CSP=="Reject CSP"))
@

>>>>

<<Fitted_and_Residual_Graph_Functions>>=
fittedGraphNoCSP <- function(data){
  fittedGraphData <- select(data, Year, Country, iGDP, fitted, fittedExp, flavor, energy, CSP, nest)
  fittedGraphData <- melt(fittedGraphData, 
                     measure.vars = c("iGDP", "fitted", "fittedExp"), 
                     value.name = "value", 
                     variable.name = "variable")
  fittedGraphData$label <- with(fittedGraphData, paste(variable, energy))
  # Now collapse some of the information. Don't need multiple versions of the exponential model, e.g.
  fittedGraphData <- filter(fittedGraphData, label=="iGDP With energy" | label=="fittedExp With energy" |
                       label == "fitted With energy" | label == "fitted Without energy")
  fittedGraphData$label <- plyr::mapvalues(fittedGraphData$label, 
                                      from = c("iGDP With energy", 
                                               "fittedExp With energy",
                                               "fitted With energy",
                                               "fitted Without energy"), 
                                      to = c("GDP", 
                                             "Reference model",
                                             "With energy",
                                             "Without energy"))
  # Relevel factors
  fittedGraphData$label <- relevelFactor(as.factor(fittedGraphData$label), 
                                         c("GDP", "Reference model", "Without energy", "With energy"))
  # Make (and return) the graph
  ggplot() +
    geom_line(data=fittedGraphData, mapping=aes(x=Year, y=value, group=label, linetype=label)) +
    scale_linetype_manual(values=c(1,4,2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

@@

The capital-labor Cobb-Douglas production function shown in Equation~\ref{eq:CD}
can be augmented to include an energy term:
%
\begin{equation} \label{eq:CDe}
  y = \theta A k^{\alpha_1} l^{\alpha_2} e^{\alpha_3} \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; ,
\end{equation}
%
where $e \equiv E/E_0$, and $E$ is in units of energy per time, typically TJ/year.
The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha_1 + \alpha_2 + \alpha_3 = 1$.
The term $A$ is known as total factor productivity,
and $\lambda$ is the Solow residual. 
%
<<Graphs_CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models that reject the cost-share theorem. Solid line is historical GDP.">>=
fittedGraphNoCSP(filter(fitted_and_resid_data, model=="CD" & CSP=="Reject CSP"))
@

%
<<Graphs_CD-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for Cobb-Douglas models.">>=
residualPlot(filter(fitted_and_resid_data, model=="CD" & CSP=="Reject CSP"))
@
